---
title: "COVID-19 Research Group: EDA and Modeling"
author: "Eddy D. Varela, Chenyang Sun, Jackson Bibbens"
date: "4/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(data.table)
countyData = read.csv("./data/countyData.csv")
#Edit Here
countyData[which.max(countyData$Income.2018),]$Pop.2019 = 8336817
countyData[which.max(countyData$Income.2018),]$Beds = 25961
countyData[which.max(countyData$Income.2018),]$popDens = 27547
countyData[which.max(countyData$Income.2018),]$Income.2018 = 76969

countyData$deathRatio<-(countyData$totalDeaths)/(countyData$Pop.2019)
countyData$caseRatio<-(countyData$totalCases)/(countyData$Pop.2019)
countyData<-countyData[which(!is.na(countyData$Beds)),]
countyData<-countyData[which(!is.na(countyData$Income.2018)),]
#Edit Here
countyData<-countyData[which(countyData$totalDeaths!=0),]
newdata<-countyData[,c("deathRatio", "caseRatio", "Income.2018","Beds","StayHomeDate","Pop.2019","popDens")]
```

There are a small number of observations for which there are not Beds or Income measurements; those are discarded. For sake of continuity, we code date as an ordered categrical variable in the form of serial date. Examining the histograms of the variables, we see that every single one except for date displays strong right skew, which suggests a logarithmic transform for each of them. Note that the +0.00001 is added to include measurements of 0, since log(0) is undefined.

```{r}
par(mfrow=c(2,2))
newdata$date<-as.numeric(as.Date(newdata$StayHomeDate))
hist(newdata$deathRatio)
hist(newdata$caseRatio)
hist(newdata$Income.2018)
hist(newdata$Beds)
hist(newdata$date)
hist(newdata$Pop.2019)
hist(newdata$popDens)

```

```{r}
newdata$Income<-as.numeric(newdata$Income)
newdata$log.Beds<-log(newdata$Beds)
newdata$log.deathRatio<-log(newdata$deathRatio+.00001)
newdata$log.caseRatio<-log(newdata$caseRatio+0.00001)
newdata$log.pop<-log(newdata$Pop.2019+0.00001)
newdata$log.popDens<-log(newdata$popDens)
newdata$log.Income<-log(newdata$Income.2018)
newdata$bedsPerCap = newdata$Beds/newdata$Pop.2019
boxplot(newdata$log.deathRatio~newdata$date,xlab="Date",ylab="log(deathRatio)")
```

Now we examine a paired scatterplot. We see quite a bit of multicollinearity between log(Beds) and population metrics, which makes sense since a large population mandates the construction of more beds. Also, some scatterplots seem to have a triangular formation; this may be attributed to the rarity of observations with extremely large values, which is made more evident by the logarithmic transform.

```{r}
vars<-with(newdata, cbind(log.deathRatio,log.caseRatio,log.Income,log.Beds,date,log.pop,log.popDens))
source("../../../panelfxns.R") 
pairs(vars,lower.panel=panel.cor)
```

We now fit the initial model with all predictors. Suspecting an interaction between population density and case ratio, we also add the interaction term.

```{r}
lm.init<-lm(log.deathRatio~log.caseRatio+log.Income+log.Beds+date+log.pop+log.popDens+log.popDens*log.caseRatio, data=newdata)
summary(lm.init)
```

```{r,echo=F}
library(MASS)
par(mfrow=c(2,2))
plot(lm.init$residuals~lm.init$fitted.values, xlab = "Fitted Values", ylab="Residuals")
qqnorm(lm.init$residuals)
```

Looking at the diagnostics, we see that the residual plot looks patternless and uniform in variance except for the line corresponding to zero deaths. The normal quantile plot looks very straight with no deviations even at the tails, normality is satisfied.

We imediately see that Income and Beds are each not significant given the other predictors, with p-values above 0.9; in fact, after removing one of them, the resulting model allows the removal of the other. We fit a model with both removed:

```{r}
lm.final<-lm(log.deathRatio~log.caseRatio+date+log.pop+Beds+log.popDens+log.popDens*log.caseRatio, data=newdata)
summary(lm.final)
anova(lm.final)
vif(lm.final)
```

We have arrived at a good candidate for a final model; all predictors and interactions also happen to be significant. We re-evaluate the diagnostics.

```{r,echo=F}
library(MASS)
par(mfrow=c(2,2))
plot(lm.final$residuals~lm.final$fitted.values, xlab = "Fitted Values", ylab="Residuals")
qqnorm(lm.final$residuals)
```

Again, aside from the line corresponding to the huge number of observations with zero deaths, the residual plot looks patternless and uniform in variance. The normal quantile plot looks very straight with no deviations even at the tails; normality is well satisfied. The final model is settled.



#Testing some alternative models
```{r moreModels}

model3 = lm(log.deathRatio~.+log.caseRatio*log.popDens, data=newdata[,-c(1,5)])
model3 = step(model3,scope =list(lower= ~1,upper = ~.), direction="backward", data=newdata[,-c(1,5)])

summary(model3)
aov(model3)
par(mfrow=c(2,2))
plot(model3$residuals~model3$fitted.values, xlab = "Fitted Values", ylab="Residuals")
qqnorm(model3$residuals)



model4 = lm(log.deathRatio~Pop.2019+popDens+date+log.Beds+log.pop+log.popDens, data=newdata[,-c(1,5)])
summary(model4)
vif(model4)
```



